---
title: "Mapping with R"
author: "Abhishek Kumar"
date: "7 September 2020"
output: github_document
---

# Data

I want to prepare a map for global distribution of tropical forests. For this,
I will be using the spatial data provided by the 
[The Nature Conservancy](https://www.nature.org/en-us/).

* [Terrestrial Ecoregions](https://geospatial.tnc.org/datasets/b1636d640ede4d6ca8f5e369f2dc368b
) spatial data [51 Mb] 

The country level data for world countries can be obtained from the [Natural Earth](https://www.naturalearthdata.com/)
at desired level of scale from the following links:

* [Cultural Admin 0 – Countries at 1:50](https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip)

# Packages

The `sf` package will be used to read the shapefiles using the \verb|st_read()|
function and the package `tmap` will be used for visualisation and mapping. The
`dplyr` package will be used for data manipulation.

```{r}
library(sf)
library(tmap)
library(dplyr)
```

# Reading the Data

The Geospatial Data Abstraction Library (GDAL) is the swiss army knife for spatial 
data: it reads and writes vector and raster data from and to practically every 
file format, or database, of significance. Package `sf` reads and writes using 
GDAL by the functions $\verb |st_read()|$ and $\verb |st_write()|$.

The data model GDAL uses:

* a data source, which may be a file, directory, or database
* a layer, which is a single geospatial dataset inside a file or directory or 
  e.g. a table in a database.
* the specification of a driver (i.e., which format)
* driver-specific reading or writing data sources, or layers

This may sound complex, but it is needed to map to over 200 data formats! Package 
`sf` tries hard to simplify this where possible (e.g. a file contains a single 
layer), but this vignette will try to point you to the options

```{r}
ec_sf <- st_read("~/ArcGIS/Ecoregions/terr-ecoregions-TNC/tnc_terr_ecoregions.shp")
world <- st_read("~/ArcGIS/Admin/World_admin/ne_50m_admin_0_countries.shp")
```

```{r}
ec_sf %>% select(WWF_MHTNUM, WWF_MHTNAM) %>% group_by(WWF_MHTNUM, WWF_MHTNAM) %>% summarise(n = n())
```


#Processign data

The variable `WWF_MHTNUM` uses values for different ecoregions

* 1. Tropical and Subtropical Moist Broadleaf Forests
* 2. Tropical and Subtropical Dry Broadleaf Forests
* 3. Tropical and Subtropical Coniferous Forests

So, we can filter the values either <4 or <=3

```{r}
ec_df <- ec_sf %>% filter(WWF_MHTNUM < 4)
```

# Plotting the data

* The `tm_shape()` function takes a spatial object (with a class from `sf`, `sp`, 
`stars`, or `raster`). 
* `tm_fill()` used to fill the polygons

```{r}
tm_shape(world) + tm_fill(col = "grey95") + 
      tm_shape(ec_df) + tm_fill(col = "seagreen") +
      tm_layout(bg.color = "lightskyblue") +
      tm_graticules(alpha = 0.2)
```


